Multi-Core Model Checking and Maximum Satisfiability Applied to Hardware-Software Partitioning

### 2016-08-23 Docs/IJES/journal.tex
paper review

\begin{abstract}
\textcolor{blue}{Bounded Model Checking (BMC) based on Satisfiability Modulo Theories (SMT) is well known by its capability to verify software. Its use as optimization tool, to solve hardware and software (HW-SW) partitioning problem, is something new. Moreover, its integration with the maximum satisfiability solver $\nu$Z tool, which provides a portfolio of approaches for solving linear optimization problems over SMT formulas, is unprecedented. We present new alternative approaches to solve the HW-SW partitioning problem. First, we use BMC based on SMT in conjunction with a multi-core support using Open Multi-Processing to create four variants to solve the partitioning problem. In other approach, we integrate the maximum satisfiability solver $\nu$Z tool into the BMC tool, making BMC as a specialized solution for optimization in a single-core environment. Multi-core SMT-based BMC approach allows initializing many verification instances using different approaches based on the number of available processing cores, where each instance checks for a different optimum value until the optimization problem is satisfied. We implement five different algorithms on top of the Efficient SMT-Based Context-Bounded Model Checker (ESBMC). We compare all proposed approaches to a state-of-the-art optimization tool (MATLAB). Experimental results show that there is no single optimization tool to solve all HW-SW partitioning benchmarks, but based in medium complexity benchmarks, ESBMC integrated with $\nu$Z was faster and precise.}
\end{abstract}

@@

%-----------------related work came to here and was rewriten
\textcolor{blue}{Since the first decade of 2000s, two main paths have been tracked to solve the HW-SW partitioning problem, {\it i.e.}, to find the exact solution of a optimization problem, as presented by~\cite{Mann2007}; to use heuristics to speed up performance time, as presented by~\cite{Arato2003} and~\cite{Arato2005}. It is worth mentioning that the using heuristics the final solution is not necessarily an optimal global solution.}

@@

@@

Our experiments are based on a set of publicly available benchmarks. All benchmarks, tools, and results of our evaluation are available on a supplementary web page \texttt{http://esbmc.org/}.

@@

The informal model can be described by five characteristics. First, there is only one software context, {\it i.e.}, there is just one general-purpose processor, and there is only one hardware context. The components of the system must be mapped to either one of these two contexts. Second, the software implementation of a component is associated with a software cost, which is the running time of the component. Third, the hardware implementation of a component has a hardware cost, which can be area, heat dissipation, or energy consumption \textcolor{blue}{(the decision is made by user or from project requirement).} Fourth, based on the premise that hardware is significantly faster than software, the running time of the components in hardware is considered as zero. Finally, if two components are mapped to the same context, then there is no overhead of communication between them; otherwise, there is an overhead. The consequence of these assumptions is that scheduling does not need to be addressed in this work. Hardware components do not need scheduling, because the running time is assumed to be zero. Because there is only one processor, software components do not need to be scheduled as well. Therefore, the focus is only on the partitioning problem. That configuration describes a first-generation co-design, \textcolor{blue}{with a simplified HW-SW model,} where the focus is on bipartitioning, as presented by~\cite{Teich2012}.

@@

The ILP and GA were taken from our previous studies, from~\cite{Trindade2015} and~\cite{Trindade2016}. \textcolor{blue}{Both use slack variables in order to eliminate the modulus operator of Eq.~\eqref{hw-sw-partitioning} and to use commercial tools.} However, GA had improvements from the parameters of related studies in order to increase the solution accuracy without producing timeout. The tuning was performed by empirical tests and resulted in changing of three parameters, which are passed to the function \textit{ga} of MATLAB, as \textcolor{blue}{mentioned} in~\cite{TheMathWorks2013}: the population size was set from $300$ to $500$, the Elite count changed from $2$ (default value) to $50$, and the number of generations changed from $100*NumberOfVariables$ (default) to $75$. \textcolor{blue}{All ILP and GA parameters were presented and discussed in~\cite{Trindade2015} and~\cite{Trindade2016}. In particular, the ILP and GA algorithms were chosen since they were also used in other related work, as~\cite{Arato2003} and~\cite{Mann2007}, thus facilitating the comparison to the present techniques (that we propose in this paper).}

@@

\textcolor{blue}{Remembering that, as defined by formal model, there is a index of each decision variable, indicated by letter "i", which ranges from zero to the number of nodes of the problem to be solved.}

@@

\vspace{2 mm}
\textcolor{blue}{It is possible to instruct ESBMC with which type of values the variables must be evaluated. Therefore, there is a declaration to populate all decision variables $x$ with non-deterministic Boolean values. As a result, the Boolean value that is assigned to each decision variable $x_i$ is actually selected by the SMT Solver, during its solving phase, which checks once all possible combinations to yield a feasible solution, {\it e.g.}, by handling the terms in the given background theory using a decision procedure~\cite{DeMoura2008,Brummayer2009}. If this is achieved, the ASSUME directive ensures the compliance of the constraint $A.x \leq b$}.

@@

\textcolor{blue}{We have to point out that the experimental results presented here depend on the computer’s processor and memory, the version of each tool (MATLAB and ESBMC), and the partitioning problem to be solved. Additionally, any change from MATLAB to another tool or to GA/ILP libraries for different programming languages can influence the measured time of the algorithms.}

@@

Based on \textcolor{blue}{mean,} standard deviation and tolerance interval to each set of time sample, it was obtained a \textcolor{blue}{confidence level} of $91.7$\% to ESBMC (sequential, SS, PB and $\nu$Z), $95.9$\% to ESBMC-PS, and $92.0$\% to ILP and GA. A timeout condition (TO) is reached when the verification time is longer than $3600$ seconds. A memory-out (MO) occurs when the tool reaches $15$ GB of memory. \textcolor{blue}{The TO was defined based on previous empirical tests as well, where a larger TO ({\it e.g.}, $5000$ seconds) did not produce substantial differences in the experimental results.}

@@

Table~\ref{Experimental-results-of-the-complex-benchmarks} shows the experimental results using Matlab (ILP and GA) and ESBMC (ESBMC-SS, ESBMC-PS, ESBMC-PB, ESBMC-$\nu$Z) tools. \textcolor{blue}{The number of nodes, edges and the initial software cost (S0) are given to each benchmark. Hp and Sp represent the results of partitioning process for each technique, i.e., the optimized value of partitioned hardware and software. T(s) means the average time taken from the experiments. Particularly about GA there is an additional row informing how far from the exact answer is the found solution.}

@@

\textcolor{blue}{Regarding the amount of MOs, the sequential ESBMC approach has to explore all (possible) states until it finds the HW-SW partitioning solution; it starts from an extreme, where all variables are selected as software, and then incrementally tests one by one to check whether a given node will be implemented in software or in hardware. In contrast, all multi-core ESBMC approaches and ESBMC-$\nu$Z, are optimized to reach faster the HW/SW partitioning solution than the sequential one, without the need for exploring all states as the sequential ESBMC approach does. As a result, if the HW/SW partitioning problem grows in complexity, then the sequential ESBMC approach tends to easily reach MO due to the state-space explosion.}

@@

We presented five approaches to solve the HW-SW partitioning problem and compared them to other state-of-the-art techniques. Experimental results showed that for a number of nodes larger than $300$, the best solution for the HW-SW partitioning problem is ILP. \textcolor{blue}{Below that limit of nodes, the best solution turns out to be ESBMC-$\nu$Z since its execution time is 4,3 to 7,5 times faster than ILP, faster than any other ESBMC approach (until 462 times faster) and its result is precise (when compared with GA).} ESBMC-PB is a viable alternative for a number of nodes lower than $150$. GA had an intermediate result in terms of performance, but the error presented from exact solution made it not acceptable to that kind of application.

@@

\bibitem[\protect\citeauthoryear{Hong and Cai}{2010}]{Hong2012}
Hong, L. and Cai, J. (2010) 'The application guide of mixed programming between MATLAB and other programming languages', {\it Proceedings of the International Conference on Computer and Automation Engineering}, pp.185--189.

>>>>

\begin{abstract}
\textcolor{blue}{Bounded Model Checking (BMC) based on Satisfiability Modulo Theories (SMT) is well known by its capability to verify software. Its use as optimization tool, to solve hardware and software (HW-SW) partitioning problem, is something new. Moreover, its integration with the maximum satisfiability solver $\nu$Z tool, which provides a portfolio of approaches for solving linear optimization problems over SMT formulas, is unprecedented. We present new alternative approaches to solve the HW-SW partitioning problem. First, we use BMC based on SMT in conjunction with a multi-core support using Open Multi-Processing to create four variants to solve the partitioning problem. In another proposed approach, we integrate the maximum satisfiability solver $\nu$Z tool into the BMC tool, making BMC as a specialized solution for optimization in a single-core environment. Multi-core SMT-based BMC approach allows initializing many verification instances using different approaches based on the number of available processing cores, where each instance checks for a different optimum value until the optimization problem is satisfied. We implement five different algorithms on top of the Efficient SMT-Based Context-Bounded Model Checker (ESBMC). We compare all proposed approaches to a state-of-the-art optimization tool (MATLAB). Experimental results show that there is no single optimization tool to solve all HW-SW partitioning benchmarks, but based on medium complexity benchmarks, ESBMC-$\nu$Z was faster and precise.}
\end{abstract}

@@

%-----------------related work came to here and was rewriten
\textcolor{blue}{Since the first decade of 2000s, two main paths have been tracked to solve the HW-SW partitioning problem, {\it i.e.}, to find the exact solution of a optimization problem, as presented by~\cite{Mann2007}; to use heuristics to speed up performance time, as presented by~\cite{Arato2003} and~\cite{Arato2005}. It is worth mentioning that using heuristics the final solution is not necessarily an optimal global solution.}

@@

\subsection{Contributions}

@@

Our experiments are based on a set of publicly available benchmarks. \textcolor{blue}{All tools, benchmarks, and results of our evaluation are available on a supplementary web page \texttt{http://esbmc.org} and \texttt{http://esbmc.org/benchmarks}}.

@@

The informal model can be described by five characteristics. First, there is only one software context, {\it i.e.}, there is just one general-purpose processor, and there is only one hardware context. The components of the system must be mapped to either one of these two contexts. Second, the software implementation of a component is associated with a software cost, which is the running time of the component. Third, the hardware implementation of a component has a hardware cost, which can be area, heat dissipation, or energy consumption \textcolor{blue}{(the decision is typically made by the engineer or from the project requirement).} Fourth, based on the premise that hardware is significantly faster than software, the running time of the components in hardware is considered as zero. Finally, if two components are mapped to the same context, then there is no overhead of communication between them; otherwise, there is an overhead. The consequence of these assumptions is that scheduling does not need to be addressed in this work. Hardware components do not need scheduling, because the running time is assumed to be zero. Because there is only one processor, software components do not need to be scheduled as well. Therefore, the focus is only on the partitioning problem. That configuration describes a first-generation co-design, \textcolor{blue}{with a simplified HW-SW model,} where the focus is on bipartitioning, as presented by~\cite{Teich2012}.

@@

The ILP and GA were taken from our previous studies, from~\cite{Trindade2015} and~\cite{Trindade2016}. \textcolor{blue}{Both use slack variables in order to eliminate the modulus operator of Eq.~\eqref{hw-sw-partitioning} and to use commercial tools.} However, GA had improvements from the parameters of related studies in order to increase the solution accuracy without producing timeout. The tuning was performed by empirical tests and resulted in changing of three parameters, which are passed to the function \textit{ga} of MATLAB, as \textcolor{blue}{mentioned} in~\cite{TheMathWorks2013}: the population size was set from $300$ to $500$, the Elite count changed from $2$ (default value) to $50$, and the number of generations changed from $100*NumberOfVariables$ (default) to $75$. \textcolor{blue}{All ILP and GA parameters were presented and discussed in~\cite{Trindade2015} and~\cite{Trindade2016}. In particular, the ILP and GA algorithms were chosen since they were also used in another related work, as~\cite{Arato2003} and~\cite{Mann2007}, thus facilitating the comparison to the present techniques (that we propose in this paper).}

@@

\textcolor{blue}{Note that, as defined by the formal model, there is an index for each decision variable, indicated by letter ``i'', which ranges from $0$ to the number of nodes of the (particular) problem to be solved.}

@@

\vspace{2 mm}
\textcolor{blue}{It is possible to instruct ESBMC with which type of values the variables must be evaluated. Therefore, there is a declaration to populate all decision variables $x$ with non-deterministic Boolean values. As a result, the Boolean value that is assigned to each decision variable $x_i$ is actually selected by the SMT Solver, during its solving phase, which checks once all possible combinations to yield a feasible solution, {\it e.g.}, by handling the terms in the given background theory using a decision procedure~\cite{DeMoura2008,Brummayer2009}. If this is achieved, then the ASSUME directive ensures the compliance of the constraint $A.x \leq b$}.

@@

\textcolor{blue}{We have to point out that the experimental results presented here depend on the computer's processor and memory, the version of each tool (MATLAB and ESBMC), and the partitioning problem to be solved. Additionally, any change from MATLAB to another tool or to GA/ILP libraries for different programming languages can influence the measured time of the algorithms.}

@@

Based on the \textcolor{blue}{mean,} standard deviation and tolerance interval to each set of time sample, it was obtained a \textcolor{blue}{confidence level} of $91.7$\% to ESBMC (sequential, SS, PB and $\nu$Z), $95.9$\% to ESBMC-PS, and $92.0$\% to ILP and GA. A timeout condition (TO) is reached when the verification time is longer than $3600$ seconds. A memory-out (MO) occurs when the tool reaches $15$ GB of memory. \textcolor{blue}{The TO was defined based on previous empirical tests as well, where a larger TO ({\it e.g.}, $5000$ seconds) did not produce substantial differences in the experimental results.}

@@

Table~\ref{Experimental-results-of-the-complex-benchmarks} shows the experimental results using Matlab (ILP and GA) and ESBMC (ESBMC-SS, ESBMC-PS, ESBMC-PB, ESBMC-$\nu$Z) tools. \textcolor{blue}{The number of nodes, edges, and the initial software cost (S0) are given to each benchmark. Hp and Sp represent the results of partitioning process for each technique, {\it i.e.}, the optimized value of partitioned hardware and software. T(s) means the average time taken from the experiments. Particularly about GA there is an additional row informing how far from the exact answer is the found solution.}

@@

\textcolor{blue}{Regarding the amount of MOs, the sequential ESBMC approach has to explore all (possible) states until it finds the HW-SW partitioning solution; it starts from an extreme, where all variables are selected as software, and then incrementally tests one by one to check whether a given node will be implemented in software or in hardware. In contrast, all multi-core ESBMC approaches and ESBMC-$\nu$Z, are (heuristically) optimized to reach faster the HW/SW partitioning solution than the sequential one, without the need for exploring all states as the sequential ESBMC approach does. As a result, if the HW/SW partitioning problem grows in complexity, then the sequential ESBMC approach tends to easily reach MO due to the state-space explosion.}

@@

We presented five approaches to solve the HW-SW partitioning problem and compared them to other state-of-the-art techniques. Experimental results showed that for a number of nodes larger than $300$, the best solution for the HW-SW partitioning problem is ILP. \textcolor{blue}{Below that limit of nodes, the best solution turns out to be ESBMC-$\nu$Z since its execution time is 4,3 to 7,5 times faster than ILP, faster than any other ESBMC approach (until 462 times faster) and its result is precise (when compared to GA).} ESBMC-PB is a viable alternative for a number of nodes lower than $150$. GA had an intermediate result in terms of performance, but the error presented from exact solution made it not acceptable to that kind of application.

@@

\bibitem[\protect\citeauthoryear{Hong and Cai}{2010}]{Hong2010}
Hong, L. and Cai, J. (2010) 'The application guide of mixed programming between MATLAB and other programming languages', {\it Proceedings of the International Conference on Computer and Automation Engineering}, pp.185--189.
